# Descriptive statistics: Income example using ABS statistics

The data set used in this task was generated using the ABS 2016 survey weekly income bands data available at http://stat.data.abs.gov.au/Index.aspx?DataSetCode=ABS_C16_T13_LGA. 

The specific income values are a randomly generated sample which does not represent real individuals, while retaining the correct statistical properties for each demographic and the nation as a whole.


##Load libraries
```{r Libraries}
# Install packages if you have not already done so 
# install.packages("dplyr") 
# install.packages("readr")
# install.packages("ggplot2") 
# install.packages("magrittr") 
# install.packages("here")

library(dplyr)    # For data manipulation 
library(readr)    # For reading data 
library(ggplot2)  # For data visualisation 
library(magrittr) # For pipes 
library(here)     # For sensible file paths 

```

## Create Output folder if one does not already exist 
``` {r outputfolder}
if(!dir.exists("outputs")) {
  dir.create("outputs")
}

``` 


##Load data

```{r LoadDataFromCSV}
file_name <- "AusIncomeSample.csv"
income_df = read_csv(here("data", file_name))

```


##Income calculations

Convert the weekly incomes into annual incomes (in AUD).

```{r AnnualIncome}
income_df$AnnualIncome <- income_df$Income * 52 

# Alternative solution: 
income_df %<>% 
  mutate(AnnualIncome = Income * 52)

```


##Plot income distribution

Plot the distribution of annual incomes on a logarithmic x-axis scale using ggplot.

The black line shows the median value while the light grey line shows the mean value.

```{r IncomePlot}
gg <- ggplot()
gg <- gg + geom_histogram(aes(x=income_df$AnnualIncome), bins=50)
gg <- gg + scale_x_continuous(trans='log10')
gg <- gg + labs(x="Annual income (AUD)", y="Count")
gg <- gg + geom_vline(xintercept = mean(income_df$AnnualIncome), size=1.5, colour = "grey")
gg <- gg + geom_vline(xintercept = median(income_df$AnnualIncome), size=1.5, colour = "black")
ggsave(here("outputs", "AnnualIncome.png"), device = "png", width = 5, height = 4)

# output to screen
gg

```


##Descriptive statistics

Output the mean and median statistics

```{r DescriptiveStatistics}
print(paste("Mean annual income (AUD):", round(mean(income_df$AnnualIncome), 1)))
print(paste("Median annual income (AUD):", median(income_df$AnnualIncome)))

```

