---
title: "2.2.3 Continuous Uniform Data"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```

## Load relevant packages

```{r}
library(ggplot2) # for ggplot()
```

## Uniform distribution on (a, b)

```{r}
a <- 0 
b <- 100 

# Calculate the population mean 
population_mean <- (a + b)/2 
cat(paste0('Population Mean: ',population_mean))
```

```{r}
# Calculate the population variance 
population_variance <- ((b - a)^2)/12 
cat(paste0('Population Variance: ',round(population_variance, 3)))
```

```{r}
# Calculate the population standard deviation
population_sd <- abs(b - a)/sqrt(12)
cat(paste0('Population Standard Deviation: ',round(population_sd, 3)))
```

## Generate 5000 random observations on X as defined above 

```{r}
random_number <- 234 # this can be any number
set.seed(random_number) # set.seed() allows reproducibility. 
# This is useful when using random processes i.e., runif() 

n <- 5000 # sample size 
r_data <- runif(n, min = a, max = b) 
head(r_data, 10) # look at first 10 sampled values
```

\newpage
## Plot distribution of sample

Note that the histogram appears quite flat over (0, 100) as anticipated. A histogram that exactly matched the theoretical distribution would look suspicious! 

```{r}
ggplot() + 
  geom_histogram(aes(r_data), boundary = b, bins = 10, color = 'white', fill = 'steelblue') + 
  scale_x_continuous(limits = c(a, b)) + 
  labs(x = "x", y = 'Frequency', title = "Histogram of simulated uniform data on (0, 100)") 
```

## Calculate summary statistics of sample

* The sample mean is close to the population mean.
* The sample variance is close to the population variance.
* The sample standard deviation is close to the population standard deviation.


```{r}
# Calculate the sample mean  
sample_mean <- mean(r_data) 
cat(paste0('Sample Mean: ',round(sample_mean,3)))
```

```{r}
# Calculate the sample variance  
sample_var <- var(r_data) 
cat(paste0('Sample Variance: ',round(sample_var, 3)))
```

```{r}
# Calculate the sample standard deviation (s)
s <- sqrt(sample_var) 
cat(paste0('Sample Standard Deviation: ',round(s, 3)))

```


