---
title: "2.2.3 Discrete Uniform Data"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE} 
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
```

## Load relevant packages

```{r}
# Ensure you have purrr installed
library(purrr) # for rdunif() function
library(plyr) # for count() function
library(ggplot2) # for ggplot() function
```


## Discrete uniform distribution on integers from 1 to 6

```{r}
a <- 1 
b <- 6

# Calculate the population mean 
population_mean <- (a + b)/2 
cat(paste0('Population Mean: ',population_mean))
```

```{r}
# Calculate the population variance 
population_variance <- (b^2 - 1)/12 
cat(paste0('Population Variance: ',round(population_variance, 3)))
```

```{r}
# Calculate the population standard deviation
population_sd <- sqrt(population_variance)
cat(paste0('Population Standard Deviation: ',round(population_sd, 3)))
```

## Generate 5000 random observations on X as defined above 

```{r}
random_number <- 234 # this can be any number
set.seed(random_number)

n <- 5000 # sample size 
r_data <- rdunif(n, b, a)
head(r_data, 10) # look at first 10 sampled values
```

## Generate frequency table and calculate proportion 

```{r}
distn <- count(r_data,vars = 'r_data')
distn$propn <- distn$freq/n
distn
```

## Plot distribution of sample

Note that the plot appears quite flat as anticipated. A plot that exactly matched the theoretical distribution would look suspicious!

```{r}
ggplot(data=distn, aes(x=as.factor(r_data), y=propn)) +  
  geom_bar(stat="identity", width=0.5, fill="steelblue") +  ylim(0,0.2) +
  labs(x='x', y='Proportion',title='Empirical discrete uniform distribution')
```

## Calculate summary statistics of sample

* The sample mean is close to the population mean.
* The sample variance is close to the population variance.
* The sample standard deviation is close to the population standard deviation.


```{r}
# Calculate the sample mean  
sample_mean <- mean(r_data) 
cat(paste0('Sample Mean: ',round(sample_mean,3)))
```

```{r}
# Calculate the sample variance  
sample_var <- var(r_data) 
cat(paste0('Sample Variance: ',round(sample_var, 3)))
```

```{r}
# Calculate the sample standard deviation (s)
s <- sqrt(sample_var) 
cat(paste0('Sample Standard Deviation: ',round(s, 3)))

```


