---
title: "2.3.3 Binomial Data"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```

## Load relevant packages

```{r}
library(plyr) # for count() function
library(ggplot2) # for ggplot() function
```


## Binomial distribution with n = 10 and p = 0.5 

```{r}
n <- 10 
p <- 0.5 

# Calculate the population mean 
population_mean <- n*p 
cat(paste0('Population Mean: ',population_mean))
```


```{r}
# Calculate the population variance 
population_variance <- n*p*(1-p) 
cat(paste0('Population Variance: ',population_variance))
```


```{r}
# Calculate the population standard deviation 
population_standard_deviation <- sqrt(population_variance) 
cat(paste0('Population Standard Deviation: ',population_standard_deviation))
```

## Generate 5000 random observations on X as defined above 

```{r}
random_number <- 234 # this can be any number
set.seed(random_number)

nbr_x <- 5000 # number of simulated values 
r_data <- rbinom(nbr_x, n, p)
head(r_data, 10) # look at first 10 sampled values
```

## Generate frequency table and calculate proportion 

```{r}
distn <- count(r_data,vars = 'r_data')
distn$propn <- distn$freq/nbr_x
distn
```

## Add a graph of the distribution 

Note that the plot displays approximate symmetry as anticipated. A plot that exactly matched the theoretical distribution would look suspicious! 

```{r fig.height = 4, fig.width = 7}
ggplot(data=distn, aes(x=as.factor(r_data), y=propn)) +  
  geom_bar(stat="identity", width=0.5, fill="steelblue") +  
  ylim(0,max(distn$propn)+0.02) + 
  geom_text(aes(label = round(distn$propn,3), y = distn$propn), vjust = -1) +
  labs(x='x', y='Proportion',title='Empirical binomial distribution')
```

## Calculate summary statistics of sample

* The sample mean is close to the population mean.
* The sample variance is close to the population variance.
* The sample standard deviation is close to the population standard deviation.

```{r}
# Calculate the sample mean  
sample_mean <- mean(r_data) 
cat(paste0('Sample Mean: ',round(sample_mean,3)))
```

```{r}
# Calculate the sample variance  
sample_var <- var(r_data) 
cat(paste0('Sample Variance: ',round(sample_var, 3)))
```

```{r}
# Calculate the sample standard deviation (s)
s <- sqrt(sample_var) 
cat(paste0('Sample Standard Deviation: ',round(s, 3)))

```



