---
title: "2.4.3 Normal data"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```

## Load relevant packages

```{r}
library(ggplot2) # for ggplot() function
```


# Normal distribution with mean = 250 and sd = 20 

```{r}
population_mean <- 250 
population_sd <- 20 
```

## Generate 5000 random observations on X as defined above 

```{r}
random_number <- 234 # this can be any number
set.seed(random_number)

nbr_x <- 5000 # number of simulated values 
r_data <- rnorm(nbr_x, population_mean, population_sd) 
head(r_data, 10) # look at first 10 sampled values
```

## Add a graph of the distribution 

Note that the histogram looks normal as anticipated. A histogram that exactly matched the theoretical distribution would look suspicious!

```{r fig.height = 4, fig.width = 7}
ggplot() +
  geom_histogram(aes(r_data), bins = 20, color = 'white', fill = 'steelblue') +
  labs(x = "x", y = 'Frequency', title = "Histogram of simulated normal data")
```

## Calculate summary statistics of sample

* The sample mean is close to the population mean.
* The sample standard deviation is close to the population standard deviation.

```{r}
# Calculate the sample mean  
sample_mean <- mean(r_data) 
cat(paste0('Sample Mean: ',round(sample_mean,3)))
```

```{r}
# Calculate the sample standard deviation (s)
sample_var <- var(r_data) # first calculate the sample variance
s <- sqrt(sample_var) 
cat(paste0('Sample Standard Deviation: ',round(s, 3)))

```



