---
title: "2.4.1 Normal distribution"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```

## Load relevant packages

```{r}
library(ggplot2) # for ggplot()
```

```{r echo=FALSE}

## We will create a function for generating density plots

## There is a lot happening with the code below, but don't worry, 
## you will NOT be asked to replicate this user-generated function 
## in any assessments, this is simply just set up to generate plots 
## specific to this script.

plot_normal <- function(x, mu, sd, a="", b="", area = "less"){
  # Define area to highlight
  
  if (area == "less") {
    auc <- c(0,dnorm(seq(mu-sd*4,x,sd*0.01), mean = mu, sd = sd),0)
    x_values <- c(mu-sd*4,seq(mu-sd*4,x,sd*0.01),x)
  } else {
    if (area == "greater") {
      auc <- c(0,dnorm(seq(x,mu+sd*4,sd*0.01), mean = mu, sd = sd),0)
      x_values <- c(x,seq(x,mu+sd*4,sd*0.01),mu+sd*4)
    } else {
      if (area == "between") {
        auc <- c(0,dnorm(seq(a,b,sd*0.01), mean = mu, sd = sd),0)
        x_values <- c(a,seq(a,b,sd*0.01),b)
      }
    }
  }
  
  # Create probability statement
  
  if (area == "less") {
    prob_statement <- paste("Pr(X < ",x,") = ", round(pnorm(x,mu,sd),3), sep = "")
  } else {
    if (area == "greater") {
      prob_statement <- paste("Pr(X > ",x,") = ", 
                              round(pnorm(x,mu,sd,lower.tail = FALSE),3), sep = "")
    } else {
      if (area == "between") {
        prob_statement <- paste("Pr(",a," < x < ",b,") = ", 
                                round(pnorm(b,mu,sd) - pnorm(a,mu,sd),3), sep = "")}
      else {
        prob_statement <- ""
      }
    }
  }
  
  
  # Plot density
  curve(expr = dnorm(x,mu,sd), 
        xlim = c(mu-sd*4,mu+sd*4), 
        main = paste("Normal Distribution, Mean = ",mu,", Sigma = ",sd),
        ylab = "Density")
  if (area != "") {
    polygon(x = x_values, y = auc, col = "tomato")
    text(x = mu-sd*4, y = dnorm(mu-sd/4,mu,sd), labels = prob_statement, pos = 4)
  }
  
}
```

### Standard normal, Z

We call $Z$ the **standard normal**. Here is what the standard normal density looks like: 

```{r fig.width = 7, fig.height = 4}
data = data.frame(x = c(-4, 4))

ggplot(data = data, aes(x)) + 
  stat_function(fun = dnorm, geom = "area", fill = "steelblue", alpha = 0.5, 
                n = 101, args = list(mean = 0, sd = 1)) +  
  xlab("") + ylab("") 
```

### Standard normal, Z versus normal variable with mean = 1 and sd = 1 

Take $Z$ as our starting point or frame of reference.  If we increase $\mu$ from 0 to 1, while keeping $\sigma = 1$, then the shape of the blue curve doesn't change, but the whole curve shifts one unit to the right, as shown below. 

```{r fig.width = 7, fig.height = 4}
ggplot(data = data, aes(x)) + 
  stat_function(fun = dnorm, geom = "area", fill = "darkgrey", alpha = 0.5, 
                n = 101, args = list(mean = 0, sd = 1)) +  
  stat_function(fun = dnorm, geom = "area", fill = "steelblue", alpha = 0.5, 
                n = 101, args = list(mean = 1, sd = 1)) +  
  xlab("") + ylab("") 

```

### Standard normal, Z versus normal variable with mean = 0 and sd = 1.5 

Once again, take $Z$ as our starting point or frame of reference.  If we increase $\sigma$ from 1 to 1.5, while keeping $\mu = 0$, then the blue curve remains centred at zero, but its spread has increased, as shown below. 

```{r fig.width = 7, fig.height = 4}
ggplot(data = data, aes(x)) + 
  stat_function(fun = dnorm, geom = "area", fill = "darkgrey", alpha = 0.5, 
                n = 101, args = list(mean = 0, sd = 1)) +  
  stat_function(fun = dnorm, geom = "area", fill = "steelblue", alpha = 0.5, 
                n = 101, args = list(mean = 0, sd = 1.5)) +  
  xlab("") + ylab("") 
```

What is also clear from looking at all of these curves above is that a normal distribution is always symmetric and centred about its mean.  

\newpage
## The following example illustrates how to solve probability questions using pnorm() and qnorm(). 

A company that produces battery operated vacuum cleaners estimates that demand for its leading model in the following year will be normally distributed with mean = 8 million units and sd = 2 million units. Run the code below to answer the following:   

a) The probability that demand will fall below 5 million units. 
```{r fig.width = 7, fig.height = 4}
# (a) P(X <= 5)
mean <- 8  
sd <- 2 

x_a <- 5 
ans.a <- pnorm(x_a, mean, sd, lower.tail = TRUE) 
print(round(ans.a,3))

# Let's visualise this using a density plot 
plot_normal(5,8,2)
```

\newpage
b) The probability that demand will exceed 12 million units. 
```{r fig.width = 7, fig.height = 4}
# (b) P(X > 12) 

x_b <- 12 
ans.b <- pnorm(x_b, mean, sd, lower.tail = FALSE) 
print(round(ans.b,3))
 
# Let's visualise this using a density plot 
plot_normal(12,8,2,area='greater')
```

\newpage
c) The 95th percentile of the demand for the leading model. 
```{r fig.width = 7, fig.height = 4}
# (c) Find x such that  P(X <= x) = 0.95 

proportion <- 0.95  

ans.c <- qnorm(proportion, mean, sd, lower.tail = TRUE) 
print(round(ans.c,3))

# We can see what this looks through a density plot 
plot_normal(round(ans.c,3),8,2)
```

