---
title: "2.6.2 Generating and visualising data from the bivariate normal distribution"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```

## Load relevant packages

```{r}
# Ensure you have "MASS" and "VGAM" packages installed
library(MASS) # mvrnorm()
library(VGAM) # pbinorm()
library(ggplot2) # ggplot()
library(tidyr) # gather()
```


# Illustrate properties of bivariate normal using randomly generated data

```{r}
# Suppose X and Y have a bivariate normal distribution with parameter values: 

mean_X <- 0 # mean of X 
mean_Y <- 0 # mean of Y 

sd_X <- 1 # standard deviation of X 
sd_Y <- 1 # standard deviation of X 

cor_XY <- 0.5 # correlation of X and Y 

mu <- c(mean_X,mean_Y) # vector of means
cov_XY <- sd_X*sd_Y*cor_XY # rearranging the correlation formula
sigma <- matrix(c(sd_X^2, cov_XY, cov_XY, sd_Y^2), nrow = 2) # variance-covariance matrix
```

In the code that follows we perform the following sub-tasks:  

1. Generate 5000 random observations on a bivariate normal 
2. Calculate the sample statistics 
3. Produce a scatterplot of the sample data 
4. Produce the histograms for the individual X and Y variables 
5. Calculate a sample proportion of P(X < -0.5 and Y < 0.7). 

\newpage
# Sub-task 1: Generate 5000 random observations on (X, Y) 

```{r}
random_number <- 234 # this can be any number
set.seed(random_number)

n <- 5000 # number of simulated values 
r_data <- mvrnorm(n, mu, sigma) 
colnames(r_data) <- c("X", "Y") 
 
head(r_data,10) # look at first 10 sampled values
```


# Sub-task 2: Calculate the sample statistics

```{r}
colMeans(r_data) # produces sample means 

var(r_data) # produces sample variance-covariance matrix 

cor(r_data) # produces sample correlation matrix 
```

\newpage
# Sub-task 3: Produce a scatterplot of the sample data 

```{r fig.width = 7, fig.height = 4}
r_data_df <- as.data.frame(r_data)
ggplot(r_data_df, aes(x=X, y=Y)) + geom_point()
```

# Sub-task 4: X and Y both exhibit characteristics of standard normal distributions 

```{r fig.width = 7, fig.height = 4}

r_data_df <- gather(r_data_df,"key","value","X", "Y")

ggplot(r_data_df, aes(x=value)) + 
  geom_histogram(bins = 40, color = "white", fill = "steelblue") + 
  facet_grid(key ~ .) + 
  labs(x = '', y= 'Frequency', title = 'Histogram of simulated normal data')
```

# Sub-task 5: Calculate a sample proportion of P(X $\leq$ -0.5 and Y $\leq$ 0.7) 

```{r}
prob <- pbinorm(0.5, 0.5, mean_X, mean_Y, sd_X^2, sd_Y^2, cov_XY) 
round(prob,3)

propn <- sum(r_data[,1] < -0.5 & r_data[,2] < 0.7)/n
round(propn,3)
```
