---
title: "3.2.5 Applications of the Central Limit Theorem - Test Yourself (Solution)"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```

## Load relevant packages

```{r}
library(dplyr) # for %>% pipelines
```

Suppose that the probability distribution of monthly incomes of account executives has mean $20,000 and standard deviation $5,000. We have randomly selected 64 account executives, and want to find the probability that the sample mean exceeds $20,500 per month.

### Constants - We've been given these in the question 

```{r}
# Constants 
mu <- 20000   # The population mean (given) 
sigma <- 5000 # The population standard deviation (given)
n <- 64       # The sample size (given) 

x_bar <- 20500 # sample mean of interest (given)
```

### Question 1: What is the standard deviation of the sample mean? 

```{r}
se <- sigma / sqrt(n)
se %>% round(digits = 3)
```
 
### Question 2: By the CLT, the distribution for the sample mean (x-bar) will be a normal distribution, with mean given by the population mean (20000) and standard deviation given by the standard error (625). 

```{r}
z <- (x_bar - mu) / se 

# or 
# z <- (x_bar - mu) / (sigma / sqrt(n)) 

z %>% round(digits = 3) 
```

### Question 3: Therefore, what is the probability that the sample mean exceeds $20,500 per month? Giver your answer to at least 2 decimal places. 

```{r}
p_val <- pnorm(q = x_bar, mean = mu, sd = se, lower.tail = FALSE)
p_val %>% round(digits = 3)
```