---
title: "4.3.3 Applying a t-test for one sample"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```

## Load relevant packages

```{r}
library(ggplot2) # for ggplot()
library(dplyr) # for %>% pipelines
```

We have some diameter data (in cm) for a motorcycle brake disc manufacturer. The manufacturer claims their brake diameter is in keeping with the regulation diameter of 12 inches (30.48 cm) for this part. Conduct a two tailed test at 5% significance level to decide whether we should reject the manufacturer's claim. 

## Two-tailed test

First let's state the statistical hypotheses given the above information.


$$H_{0} : \mu = 30.48$$

$$H_{A} : \mu \ne 30.48$$

We will explore the data to determine whether a z-test or t-test is suitable.

```{r}
# load the data set
brakes_df <- read.csv("new_disc_brakes.csv") # make sure to set your work directory accordingly
head(brakes_df)
```

```{r}
# check how large sample is
cat(paste('Length of brakes data = ', length(brakes_df$brake_diameter)))
```

Population standard deviation was not provided so we must estimate it from sample data. Also dataset is small (n <=30). Hence we require a t-test, however due to the sample size we must check to see whether our sample data exhibits a normal distribution.

## Assess normality

```{r}
# Here are some sample statistics: 
brakes_df %>% summarise(Min = min(brake_diameter, na.rm = TRUE), 
                        Max = max(brake_diameter, na.rm = TRUE), 
                        Mean = mean(brake_diameter, na.rm = TRUE), 
                        SD = sd(brake_diameter, na.rm = TRUE), 
                        n = n(), 
                        Missing = sum(is.na(brake_diameter))) 
```


```{r fig.width = 7, fig.height = 4}
# Plot histogram to see if you can reasonably assume normality 
ggplot() +
  geom_histogram(aes(brakes_df$brake_diameter), bins = 7, color = 'white', fill = 'steelblue') +
  labs(x = "x", y = 'Frequency', title = "Histogram of brake disc diameter (cm)") + 
  geom_vline(xintercept=30.48, color = "red")
```
Histogram shape suggests you can assume normality.

\newpage

## Two-tailed t-test

Now that we can assume normality, we can perform a t-test.

```{r}
# Full output includes 95% confidence interval by default 
# If you want another level of confidence, than add conf.level as argument 
t.test(brakes_df$brake_diameter, alternative = "two.sided",mu = 30.48) 
```

```{r}
# You can also get the p-value directly via 
t.test(brakes_df$brake_diameter, alternative = "two.sided", mu = 30.48)$p.value 
```

## Interpretation: 

Since p-value (0.017) < significance level (0.05) you reject the manufacturer's claim.
