---
title: "4.3.2 Applying a two-tailed z-test for one sample"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```

## Libraries
```{r}
library(ggplot2) # for ggplot()
```


## Two tailed test

The aim of this example is to test, at a 1% significance level, the hypothesis that the average speed of a machine is **different** from 1,000 rpm. This requires a two-tailed test.

$$H_{0} : \mu = 1000$$

$$H_{A} : \mu \ne 1000$$

Let's first load the data and visualise the sample mean alongside the assumed population mean.

```{r}
# load the machine rpm data set
machine_df <- read.csv("machine_rpm.csv") # make sure to set your work directory accordingly
head(machine_df)
```

```{r}
# check how large sample is
cat(paste('Length of RPM data = ', length(machine_df$rpm)))
```

```{r fig.width = 7, fig.height = 4}
# visualise sample mean

qplot(machine_df$hours, machine_df$rpm, xlab = 'Hours', ylab='rpm') + 
  geom_hline(aes(yintercept=1000, linetype = 'Population Mean'), color = "red") + 
  geom_hline(aes(yintercept=mean(machine_df$rpm), linetype = 'Sample Mean'), color = "blue") + 
  scale_linetype_manual(name = "", values = c(2, 1), 
                      guide = guide_legend(override.aes = list(color = c("red", "blue"))))
```

Since the population standard deviation is known and the sample size is large, a z-test is conducted. To perform a z-test we will need to calculate the z-score for the sample mean.

```{r}
population_sd <- sd(machine_df$rpm)
population_mu <- 1000

# length of sample
n <- length(machine_df$rpm)

# calculate sample mean
x_bar <- mean(machine_df$rpm)
cat(paste("Sample mean: ", round(x_bar,2)))

# calculate standard error
SE <- population_sd / sqrt(n)
cat(paste("Standard error: ", round(SE,2)))

# calculate the z score
z_score <- (x_bar - population_mu) / SE
cat(paste("z score: ", round(z_score,2)))
```
 
\newpage
Now we have calculated the z-score, we can use the pnorm() function to generate a p-value. We are interested in whether the average rpm is above or below 1,000. 
 
```{r}
p_value = 2*pnorm(abs(z_score), lower.tail = FALSE)
cat(paste("p-value: ", p_value))
```

Based on the p-value and the significance level (0.01), we can conclude our findings.

```{r}
signif_level <- 0.01

statement = c("the null hypothesis at the 1% significance level.",
              "\nConclude that the average speed")

true = c("Reject", statement, "differs from 1,000 rpm.")
false = c("Fail to reject",statement, "is 1,000 rpm.")

if (p_value < signif_level) cat(paste(true)) else cat(paste(false))
```
