---
title: "4.3.2 Applying a one tailed z-test for one sample"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```

## One tailed test

The aim of this example is to test, at a 5% significance level, the hypothesis that people working in profession 7 have an IQ **higher** than the average. 

$$H_{0} : \mu = \mu_{0}$$

$$H_{A} : \mu > \mu_{0}$$

This requires an upper tailed test. IQ test score is calculated based on a normed group with an average of 100 and a standard deviation of 15. As we know the population mean to be 100, we can also write the statistical hypotheses as follows:

$$H_{0} : \mu = 100$$

$$H_{A} : \mu > 100$$


## Upper tailed test for testing differences in IQ test score

Let's first load the data and take a look at people working in profession 7.

```{r}
# load the IQ data set
iq_df <- read.csv("IQ.csv") # make sure to set your work directory accordingly
head(iq_df)
```


\newpage
Let's look at profession 7 in this example

```{r}
profession7 <- iq_df[iq_df$Profession == 7,]
head(profession7)
```
 
For IQ scores we already know the standard deviation by definition, but what about the number of samples?

```{r}
length(profession7$IQ)
```

Since the population standard deviation is known and the sample size is large, a z-test is conducted. To perform a z-test we will need to calculate the z-score for the sample mean.

```{r}
population_sd <- 15
population_mu <- 100

# length of sample
n <- length(profession7$IQ)

# calculate sample mean IQ for people working in profession 7
x_bar <- mean(profession7$IQ)
cat(paste("Sample mean: ", round(x_bar,2)))

# calculate standard error
SE <- population_sd / sqrt(n)
cat(paste("Standard deviation: ", round(SE,2)))

# calculate the z score
z_score <- (x_bar - population_mu) / SE
cat(paste("z score: ", round(z_score,2)))
```

\newpage
Now we have calculated the z-score, we can use the pnorm() function to generate a p-value.

```{r}
# Upper tail test has p-value = P(Z>z) 

signif_level <- 0.05 

p_value = pnorm(z_score, lower.tail = FALSE) 
cat(paste("p-value: ", p_value))
```

Based on the p-value and the significance level (0.05), we can conclude our findings.

```{r}
signif_level <- 0.05 

statement = c("the null hypothesis at the 5% significance level.",
              "\nConclude that those working in profession 7 have")

true = c("Reject", statement, "\na higher average IQ than the overall population.")
false = c("Fail to reject",statement, "\nthe same IQ as the overall population.")

if (p_value < signif_level) cat(paste(true)) else cat(paste(false))
```