---
title: "5.1.2: One-way (between-groups) ANOVA in R"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```

One-way ANOVA is a parametric test designed to compare the means of two or more groups. The null hypothesis states that the means of all groups to be tested are equal. As usual, the test will return a p-value in the end, and you will be able to decide whether or not to reject the null hypothesis depending on this p-value.

Assumptions are:

* independence of observations: each individual is represented by 1 entry/measurement ONLY
* normality of distribution: to be tested for each group, for example with the [Shapiro-Wilk test](https://biostats.w.uib.no/test-for-normality-shapiro-wilks-test/)
* homogeneity of variance: to be tested with, for example, [Levene's test](https://biostats.w.uib.no/test-for-homogeneity-of-variances-levenes-test/).

Here, let's suppose that you want to check whether the average size of blue ground beetles (Carabus intricatus) differs depending on their location. We consider 3 different locations, for example 3 forests beautifully named A, B and C. In each location, we measure the size (in millimeters) of 10 individuals.

## Create data

To create the corresponding dataframe in R, use the following code:

```{r create_data}
size <- c(25, 22, 28, 24, 26, 24, 22, 21, 23, 25, 26, 
          30, 25, 24, 21, 27, 28, 23, 25, 24, 20, 22, 
          24, 23, 22, 24, 20, 19, 21, 22) 

location <- as.factor(c(rep("ForestA", 10), 
                        rep("ForestB", 10), 
                        rep("ForestC", 10)))

my.dataframe <- data.frame(size, location)
my.dataframe
```

It is always nice and useful to get an overview of the whole dataset, so let’s plot the data:

```{r plot_data}
plot(size ~ location, data = my.dataframe)
```

## Test assumptions

Now we need to check the assumptions of normality of distribution and homogeneity of variance. We thus run the Shapiro-Wilk test on each group and then Levene's test. For which you will need to load/activate the package ``car`` via the command ``library(car)``.

```{r test_assumptions}
library(car)

shapiro.test(my.dataframe$size[location == "ForestA"])
shapiro.test(my.dataframe$size[location == "ForestB"])
shapiro.test(my.dataframe$size[location == "ForestC"])

leveneTest(size ~ location, data = my.dataframe, center = mean)

```

The Shapiro-Wilk test of normality tests the Null Hypothesis that the data are drawn from a normal distribution, versus the Alternative Hypothesis that they are drawn from some other (unspecified) distribution. If the p-value is greater than 0.05, there is insufficient evidence to conclude that the data are drawn from some other distribution and therefore you may assume normality - although normality is not actually proven, only assumed, as you never prove a null hypothesis. 

Therefore, each of the 3 groups (ForestA, ForestB and ForestC) are assumed to come from a normal distribution since the p-value of the Shapiro-Wilk test is greater than 0.05; additionally, variances are not different according to Levene’s test (p-value greater than 0.05).

Note: if you are a bit confused about the way data/groups are retrieved for running the Shapiro-Wilk test, here is a quick explanation. Let's consider the group ForestA: we need to tell the function to retrieve all ``size`` data located in the object ``my.dataframe`` (hence ``my.dataframe$size``) but we need to restrict the selection to data matching the criteria ForestA only (hence ``[location == ForestA]``). Putting everything together, we write ``my.dataframe$size[location == "ForestA"``] inside ``shapiro.test()``.

## Run ANOVA

Using R we can quickly perform an ANOVA test using the function ``aov``.

To carry out a one-way ANOVA use ``aov(dependent ~ independent)``, give the ANOVA model a name e.g. ``results`` and use ``summary()`` to see the output.

```{r anova}
results <- aov(size ~ location, data = my.dataframe)
summary(results)
```

This output gives the value F of the statistic (here F = 7.11) and the p-value (0.00331) which is very close to the results obtained with lm(), at least in this example. Here, the ANOVA test tells you that the null hypothesis is to be rejected, so you conclude that there exists a significant difference between some of the groups.

But this does not tell us anything about which groups have significantly different means ...

For that you need to run a post hoc test which is covered in the next task.

## Final note...

What would you do if the previously determined assumption of normality was not met?

In this case you can simply apply the non-parametric Kruskal-Wallis test, which doesn't assume normality.

```{r Kruskal}
kruskal.test(size ~ location, data = my.dataframe)
```

This test shows that the null hypothesis can be rejected. You can therefore conclude that there are differences between the group means.
