---
title: "MATH2406 Applied Analytics"
author: "Student name and number comes here"
subtitle: 'Assessment 1: Report'
output:
  pdf_document: default
  html_document:
    df_print: paged
  html_notebook: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```

**DELETE the instructional text provided in this template.** 

If you have any questions regarding the assignment instructions or the R Markdown template, please post them on the discussion board.


## **Setup**

Install and load the packages you need to produce the report here:

```{r, echo = TRUE, warnings = FALSE}

# This is a chunk where you can load the packages required for producing the report 

# Here are some example packages. You may add others if required  


library(dplyr) # Useful for data manipulation
library(ggplot2) # Useful for building data visualisations
library(knitr) # Useful for creating nice tables

```



## **Instructions** 

Follow the instructions given in the Assessment brief to fill out the template below. Remember to include R codes and outputs, and where applicable plain text explanations.


## **Task 1**

Write your plain text here. 

```{r}

# Provide your R code here


```


## **Task 2**

Write your plain text here. 

```{r}

# Provide your R code here


```


## **Task 3**

Write your plain text here. 

```{r}

# Provide your R code here


```


## **Task 4**

Write your plain text here. 

```{r}

# Provide your R code here


```


## **Task 5**

Write your plain text here. 

```{r}

# Provide your R code here


```


## **Task 6**

Write your plain text here. 

```{r}

# Provide your R code here


```


### **IMPORTANT NOTE** 

The report must be uploaded to Assignment 1 section in Canvas as a **PDF** document with R codes and outputs showing. The easiest way to do this is to:
<br>
1) Run all R chunks
<br>
2) **Preview** your notebook in **HTML** (by clicking Preview Notebook)
<br>
3) **Open in Browser (Chrome)**
<br>
4) **Right Click on the report in Chrome**
<br>
5)  **Click Print** and Select the Destination Option to **Save as PDF**. 
<br>
6) Now upload this PDF report as one single file via the Assignment 1 page in Canvas.
<br>
<br>
**Remember to DELETE the instructional text provided in the template. Failure to do this will INCREASE the SIMILARITY INDEX reported in TURNITIN** 
<br>

If you have any questions regarding the assignment instructions or the R Markdown template, please post them on the discussion board. 

<br>
