---
title: "3.2.2 Using R to demonstrate CLT"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```

## Load relevant packages

```{r}
library(ggplot2) # for ggplot()
```

```{r echo=FALSE}

## We will create a function for generating uniform density plots

## There is a lot happening with the code below, but don't worry, 
## you will NOT be asked to replicate this user-generated function 
## in any assessments, this is simply just set up to generate plots 
## specific to this script.

plotUnif <- function (min, max, type = "b", col = "black") 
{
  if (!is.numeric(min) | !is.vector(min) | any(!is.finite(min))) {
    stop("The minimun of the uniform distribution, 'min', must be a single number")
  }
  if (length(min) != 1) {
    stop("The minimun of the uniform distribution, 'min', must be a single number")
  }
  if (!is.numeric(max) | !is.vector(max) | any(!is.finite(max))) {
    stop("The maximun of the uniform distribution, 'max', must be a single number")
  }
  if (length(max) != 1) {
    stop("The maximun of the uniform distribution, 'max', must be a single number")
  }
  if ((max - min) <= 0) 
    stop("The parameters min and max represent the minimum and maximum values of the uniform distribution, respectively")
  if (!type %in% c("b", "dis", "den", "q")) 
    stop("The argument 'type' must be 'b', 'dis', 'den' or 'q'")
  if (length(col) != 1) 
    stop("The argument 'col' must be a single colour")
  if (col %in% c(NA, NaN, Inf, -Inf)) 
    stop("The argument 'col' must be a single colour")
  oldpar <- par(no.readonly = TRUE)
  on.exit(par(oldpar))
  dom = (max - min)/10
  x = seq(min - dom, max + dom, by = 0.001)
  fx = dunif(x, min = min, max = max)
  Fx = punif(x, min = min, max = max)
  Finvx = qunif(seq(0, 1, by = 0.01), min = min, max = max)
  if (type == "b") {
    par(mfrow = c(1, 3))
    plot(x, fx, type = "l", main = "Density Function", ylab = "f(x)", 
         lwd = 2, col = col)
    segments(min, 0, min, 1/(max - min), lty = 2, col = "white", 
             lwd = 3)
    segments(max, 0, max, 1/(max - min), lty = 2, col = "white", 
             lwd = 3)
    plot(x, Fx, type = "l", main = "Distribution Function", 
         ylab = "F(x)", lwd = 2, col = col)
    abline(h = c(0, 1), lty = 2, col = "gray")
    plot(seq(0, 1, by = 0.01), Finvx, type = "l", xlab = expression(tau), 
         ylab = "", main = "Quantile Function", lwd = 2, 
         col = col)
    title(ylab = expression(paste("F"^"-1", (tau), sep = "")), 
          line = 2.5, cex.lab = 1)
  }
  else if (type == "dis") {
    plot(x, Fx, type = "l", main = "Distribution Function", 
         ylab = "f(x)", lwd = 2, col = col)
    abline(h = c(0, 1), lty = 2, col = "gray")
  }
  else if (type == "den") {
    plot(x, fx, type = "l", main = "Density Function", ylab = "f(x)", 
         lwd = 2, col = col)
    segments(min, 0, min, 1/(max - min), lty = 2, col = "white", 
             lwd = 3)
    segments(max, 0, max, 1/(max - min), lty = 2, col = "white", 
             lwd = 3)
  }
  else if (type == "q") {
    plot(seq(0, 1, by = 0.01), Finvx, type = "l", xlab = expression(tau), 
         ylab = "", main = "Quantile Function", lwd = 2, 
         col = col)
    title(ylab = expression(paste("F"^"-1", (tau), sep = "")), 
          line = 2.5, cex.lab = 1)
  }
}
```

## Example 1

### CLT applied to Poisson data drawn from Poisson random variable with parameter lambda = 3. 

```{r fig.width = 7, fig.height = 4}
random_number <- 234 # this can be any number
set.seed(random_number)

lambda <- 3 # Poisson parameter value 
n_range <- 0:20

# calculate probabilities for 21 outcomes
PMF <- dpois(x = 0:20,lambda = lambda)
PMF

# create a dataframe to pass the ggplot() function
PMF_df <- data.frame(n_range,PMF)

ggplot(PMF_df, aes(x = as.factor(n_range), y = PMF)) +
geom_bar(stat="identity", width=0.5, fill="steelblue") + ylim(0,max(PMF)+0.02) +
labs(title = paste("Poisson distn, lambda = ",lambda),
x = "x",
y = "Probability Mass Function")
```

### Generate vector of 50,000 observations on Pois(3) random variable  

```{r fig.width = 7, fig.height = 4}
set.seed(random_number)

n <- 50 # sample size 
rows <- 1000 # samples 
x <- rpois(n*rows, lambda)  

head(x,10) # look at first 10 sampled values

# Let's have a look at the frequencies for each x
x_table <- as.data.frame(table(x))
x_table

# Some 'x' values were not sampled i.e., 13, 14 and 15, 
# we add dummy values to our table to improve the readability of the plot
missing_x_ticks <- data.frame('x' = c('13','14','15'), 'Freq' = rep(0,3))
x_table <- rbind(x_table[0:13,],missing_x_ticks, tail(x_table,1))
x_table$x <- factor(x_table$x, levels=x_table$x)

ggplot(x_table, aes(x = x, y = Freq)) +
geom_bar(stat="identity", width=0.5, fill="steelblue") +
labs(title = paste("Distribution of 50,000 observations on Pois(3) random variable")
     ,y = "Frequency")
```

Lastly, we generate the distribution of sample means with an overlaid normal curve.

```{r fig.width = 7, fig.height = 4}

# reshape vector into matrix with 1000 rows  
m <- matrix(x, rows)  
x.bars <- apply(m, 1, mean) # 1 returns a vector of row means  
avg.x.bars <- format(round(mean(x.bars), 3), nsmall = 3)
cat(paste('Mean of sampling distribution: ',avg.x.bars))

sd.x.bars <-format(round(sd(x.bars), 3), nsmall = 3) 
cat(paste('Standard deviation of sampling distribution: ',sd.x.bars))

x.bars <- as.data.frame(x.bars) # required for ggplot 
head(x.bars$x.bars,10) # look at first 10 sampled values 

ggplot(data = x.bars) +  
  geom_histogram(mapping = aes(x = x.bars, y=..density..), 
                 color = 'white',fill="steelblue", binwidth = 1/10) + 
  labs(title = paste("Distribution of sample means with normal curve"),
       x ="Sample mean values", y = "Density") + 
  stat_function(fun = dnorm, args = list(mean = lambda, sd = sqrt(lambda/n))) 
```

```{r}
# We can calculate the standard error and compare with the 
# standard deviation of the sampling distribution
mean.x.bars.CLT <- lambda 
sd.x.bars.CLT <- format(round(sqrt(lambda/n), 3), nsmall = 3)
cat(paste('Standard error: ',sd.x.bars))
```

\newpage
## Example 2
### CLT applied to Uniform (0, 1) data - which has mean 1/2 and variance 1/12.

```{r fig.width = 7, fig.height = 4}
set.seed(random_number)

a <- 0
b <- 1

# define x-axis 
x <- seq(a, b, length=1000)
head(x, 10)

# calculate uniform distribution 
y <- dunif(x, min = a, max = b)

# plot uniform distribution 
plotUnif(min = a, max = b, type = "den", col = "red")
```

### Generate vector of 50,000 observations Uniform(0, 1) random variable

```{r fig.width = 7, fig.height = 4}
set.seed(random_number)
n <- 50 # sample size 
rows <- 1000 # samples 

x <- runif(n*rows)  
head(x,10) # look at first 10 sampled values

ggplot() +
geom_histogram(aes(x), boundary = b, bins = 20, color = "white", fill = "steelblue") +
scale_x_continuous(limits = c(a, b)) +
labs(x = "x", y = "Frequency", title = "Histogram of x")
```

Lastly, we generate the distribution of sample means with an overlaid normal curve.

```{r fig.width = 7, fig.height = 4}

# reshape vector into matrix with 1000 rows  
m <- matrix(x, rows)  
x.bars <- apply(m, 1, mean) # 1 returns a vector of row means  
avg.x.bars <- format(round(mean(x.bars), 3), nsmall = 3)
cat(paste('Mean of sampling distribution: ',avg.x.bars))

sd.x.bars <-format(round(sd(x.bars), 3), nsmall = 3) 
cat(paste('Standard deviation of sampling distribution: ',sd.x.bars))

x.bars <- as.data.frame(x.bars) # required for ggplot 
head(x.bars$x.bars,10) # look at first 10 sampled values 

ggplot(data = x.bars) +  
  geom_histogram(mapping = aes(x = x.bars, y=..density..),  
                 color = 'white',fill="steelblue", binwidth = 1/100) +
  labs(title = paste("Distribution of sample means with normal curve"),
       x ="Sample mean values", y = "Density") +
  stat_function(fun = dnorm, args = list(mean = 0.5, sd = sqrt((1/12)/n))) 
```

```{r}
# We can calculate the standard error and compare with the 
# standard deviation of the sampling distribution
mean.x.bars.CLT <- lambda 
sd.x.bars.CLT <- format(round(sqrt(lambda/n), 3), nsmall = 3)
cat(paste('Standard error: ',sd.x.bars))
```